A Data-Driven Framework for Neural Field Modelling

### 2010-11-18 ltx/first_order_framework/LaTeX_Manuscript.tex
some formatting for the graph and tables

\begin{document}
% Title must be 150 characters or less
\title[A Data-Driven Framework for Neural Field Modelling]{A Data-Driven Framework for Neural Field Modelling}

\author{Dean R. Freestone$^{1,2,3}$, 
Parham Aram$^{4}$, 
Michael Dewar$^{3,5}$,
Kenneth Scerri$^{6}$,
David B. Grayden$^{1,2}$,
Visakan Kadirkamanathan$^{4}$}

\address{$^1$ Department of Electrical and Electronic Engineering, University of Melbourne, Melbourne, VIC, Australia
\\
$^2$ The Bionic Ear Institute, East Melbourne, VIC, Australia
\\
$^3$ Institute for Adaptive and Neural Computation, University of Edinburgh, Edinburgh, UK
\\
$^4$ Department of Automatic Control and Systems Engineering, University of Sheffield, Sheffield, UK
\\
$^5$ Department of Applied Physics and Applied Mathematics, Columbia University, New York, NY, USA
\\
$^6$ Department of Systems and Control Engineering, University of Malta, Msida, MSD, Malta}

\ead{dfreestone@bionicear.org}

\begin{abstract}
This paper presents a framework for creating neural field models from electrophysiological data. The Wilson and Cowan or Amari style neural field equations are used to form a parametric model, where the parameters are estimated from data. To illustrate the estimation framework, data is generated using the neural field equations incorporating modelled sensors so a comparison can be made between estimated and true parameters. To facilitate state and parameter estimation, we introduce a method to reduce the continuum neural field model using a basis function decomposition to form a finite-dimensional state-space model. Spatial frequency analysis methods are introduced that systematically specify the basis function configuration required to capture the dominant characteristics of the neural field. The estimation procedure consists of a two-stage iterative algorithm incorporating the unscented Rauch-Tung-Striebel smoother for state estimation and a least squares algorithm for parameter estimation. The results show that it is theoretically possible to reconstruct the neural field and estimate intracortical connectivity \parham{kernel} and synaptic dynamics with the proposed framework. %The results also illustrate a systematic loss of high spatial frequency information as the cost incurred by the model reduction procedure. 
\end{abstract}

\maketitle

@@

	 \int_\Omega \boldsymbol{\psi}^{\top} (\mathbf{r}-\mathbf{r}')f(\boldsymbol{\phi}^{\top}(\mathbf{r}')\mathbf{x}_t) \nonumber \\
	&\times& d\mathbf{r}' d\mathbf{r} \boldsymbol{\theta}  
	 + \xi\mathbf{x}_t + \boldsymbol{\Gamma}^{-1} \int_\Omega{\boldsymbol{\phi}(\mathbf{r}) e_t(\mathbf{r}) \, d\mathbf{r}}.
	 % \end{split}
\end{eqnarray}

@@

Histograms of the parameter estimates obtained over the 150 realisations are shown in Figure~\ref{fig:Figure7}. In each of the subplots of the figure the true parameter value is shown by the solid line and the mean parameter estimate is shown by the dashed line. The actual parameter values, the mean parameter estimates, standard deviations, and biases for this parameter combination are given as the first entry in Table~\ref{tab:MexicanHatKernelResults}, where other entries in the table show results from Experiment II. The biases show the percentage differences between the actual values and the mean parameter estimates. For this parameter combination the mean estimates are in good agreement (within 1 standard deviation) with the actual values of the connectivity kernel parameters. However, the time constant parameter, $\xi$, is outside the estimate distribution (with bias of $2.56\%$), with a small variance in parameter estimates. We attribute this small bias to the disparity between the actual field and the smoother estimated field from the decomposition.

The accuracy of the state estimates was evaluated by comparing the field reconstruction to the true field using the mean (over iterations) of the root mean square error (MRMSE) over space. The MRMSE of the field estimates over time for 150 realisations is shown in Figure~\ref{fig:Figure9} with the $95~\%$ confidence interval. At each time point, the error is consistently around $0.5$~mV, which is approximately $10.73~\%$ of the field voltage range ([-2.20, 2.46]~mV). The error in the reconstructed field is mostly due to the loss of high spatial frequency information from the basis function decomposition as illustrated in Figure~\ref{fig:Figure10}. The figure shows a snapshot at a single time instant of the true neural field, the estimated field, and the corresponding error, which is the difference between the true and estimated neural fields. It can be seen in Figure~\ref{fig:Figure10}(B) that the estimated field contains the basic structure of the true field, shown in Figure~\ref{fig:Figure10}(A), but is missing the higher-frequency detail, which can be see in the error shown in Figure~\ref{fig:Figure10}(C).

@@

% The size of the squares in the figure show the magnitude of the log of the error. The squares are black if the log of the error is negative (the RMSE is $<1$) and the squares are white if the log of the error is positive (the RMSE is $>1$). The largest black squares indicate the best performance and the largest white squares indicate the worst performance for the parameter combinations (\dean{need some thing more quantitative}). 
From Figure~\ref{fig:Figure12} it can be seen that the estimation accuracy depends on the parameters of the sigmoid function. This is expected since that by altering the parameters we are changing the nonlinearity and the saturation region of the sigmoid.  
\dean{PLEASE HELP TO FILL THIS SECTION IN. ANYTHING WILL DO.}
..\dean{This next bit is CRAP (please help):} When the threshold is negative or the slope is very gentle, all of the activity is being passed through the sigmoid function. However, as the slope is increased or the threshold is increased, more of the field is within the saturating region of the activation function. This is illustrated in Figure~\ref{fig:sigmoidfunctions}, where the sigmoids from the bottom left corner through to near top right corner of Figure~\ref{fig:sigmoidfunctions} are plotted. This appears to have a detrimental effect on the estimation accuracy.

\subsubsection{Experiment IV: Results for Varying the Sensor Kernel Width}
The accuracy of the estimation algorithm was tested for different observation kernels by varying the width parameter, $\sigma_m$. Figure~\ref{fig:Figure13} shows the RMSE of the connectivity kernel estimates for the various observation kernel widths. As the theory presented earlier in Section~\ref{SpectralAnalysisSection} suggests, the observation kernel acts as a spatial filter. When the observation kernel is too narrow, a small amount of high spatial frequency information is still present but not adequately sampled by the observation function, since the field is not exactly band-limited. Therefore, a small amount of aliasing occurs which results in the estimation error. When the observation kernel width is increased, the high spatial frequency information that is not adequately sampled is filtered more strongly by the observation kernel and the error decreases. However, as expected when the observation kernel becomes too large, important spatial frequency information is lost and the error becomes larger.

@@

\section{Conclusion}
\dean{I MOVED THIS FROM THE START OF THE DISCUSSION. ANY OBJECTIONS?} This paper has presented a framework for creating data-driven neural field models by developing a method for estimating intracortical connectivity and synaptic dynamics from electrophysiological data. In doing so, we have highlighted important aspects of the experiment design process, demonstrating the relationship between sensor bandwidth and the complexity of the field that can be inferred. We have derived a finite-order state-space representation of the neural field equations that allows inference to be performed efficiently. In addition, the mathematical formulation that we have presented scales well with the number of sensors. The major contribution of this framework is to formally bridge a gap between experimental and theoretical neuroscience in order to gain a better understanding of the mesoscopic dynamics of the brain.

\section{Acknowledgments}
Dean Freestone would like to thank Dr. Mark van Rossum for hosting him while visiting the University of Edinburgh where this work was undertaken and Dr. Matthias Hennig for useful discussions and feedback. This research was funded in part by the Australian Research Council (Linkage Project LP0560684). The Bionic Ear Institute acknowledges the support it receives from the Victorian State Government through the Operational Infrastructure Support Program. Dean Freestone would like to thank the University of Melbourne's Scholarships Office for the Postgraduate Overseas Research Experience Scholarship and the Harold Mitchell Foundation for a traveling scholarship for supporting this research. Parham Aram would also like to thank Andrew Hills for useful discussions and assistance with code and the generation of figures.

% This research was funded in part by the Australian Research Council (Linkage Project LP0560684). The Bionic Ear Institute acknowledges the support it receives from the Victorian State Government through the Operational Infrastructure Support Program. Dean Freestone would like to thank the University of Melbourne's Scholarships Office for the Postgraduate Overseas Research Experience Scholarship and the Harold Mitchell Foundation for a traveling scholarship and supporting this research.

\section*{References}
\bibliographystyle{unsrt} 
\bibliography{BrainIDE}

\section*{Figure Legends}
%\begin{figure}[!ht]
%\begin{center}
%%\includegraphics[width=4in]{figure_name.2.eps}
%\end{center}
%\caption{
%{\bf Bold the first sentence.}  Rest of figure 2  caption.  Caption 
%should be left justified, as specified by the options to the caption 
%package.
%}
%\label{Figure_label}
%\end{figure}

\begin{figure}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/Figure1.pdf} 
\end{center}
\caption{{\bf Mexican-hat connectivity kernel.} The function $w(\cdot)$ is rotationally symmetric (isotropic) about zero, hence a cross-section captures the important aspect of the connectivity kernel's shape. The kernel decays asymptotically to zero.}
\label{fig:Figure1}
\end{figure}

\begin{figure}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/Figure2.pdf} 
\end{center}
\caption{{\bf Example of a one-dimensional field decomposition.} A continuous, one-dimensional field decomposed by a finite number of basis functions scaled by the state vector. The dashed line depicts the field, the black solid lines shows the Gaussian basis functions and the vertical lines show the position and amplitude of the states.} 
\label{fig:Figure2}
\end{figure}

\begin{figure}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/Figure3.pdf}
\end{center}
\caption{{\bf Example of the spatiotemporal properties of the model and synthetically generated data.} \textbf{(A)} The neural field at a single time instant. (\textbf{B}) Data generated by the full neural field model from the first 5 sensors. (\textbf{C}) Power spectral density of the data generated by an observation showing the typical $1/f^2$ characteristic of intracranial EEG.} 
\label{fig:Figure3}
\end{figure}

\begin{figure}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/Figure4.pdf}
\end{center}
\caption{{\bf Spatial frequency analysis of the neural field}. (\textbf{A}) The average (over time) power in dB of the spatial frequency of the neural field. (\textbf{B}) The average (over time) power in dB of the spatial frequency of the reconstructed neural field from the basis function decomposition. The reconstructed field shows a narrower bandwidth as a result of the band limiting properties of the sensors and basis functions.}
\label{fig:Figure4}
\end{figure}

\begin{figure*}[!ht] 
\begin{center}
\includegraphics{./Graph/pdf/Figure5.pdf}	
\end{center}
\caption{{\bf Spatial frequency analysis}. The dashed line shows the cutoff frequency (-3 dB point). (\textbf{A}) The average (over time) power in dB of the spatial frequency of the observations. (\textbf{B}) The spatial frequency response of a one-dimensional model sensor. The symmetry of the sensors and field basis functions allow for the one-dimensional representation of the spectral characteristics. (\textbf{C}) The spatial frequency response of a one-dimensional neural field basis function, illustrating a narrower band width than the sensors. }
\label{fig:Figure5}
\end{figure*}

\begin{figure}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/Figure6.pdf}
\end{center}
\caption{{\bf Convergence of parameters.} The mean of the absolute error between true and estimated parameters averaged over 150 realisations. All parameters converge after six iterations, where the mean difference between errors in consecutive iterations falls below $10^{-4}$.}
\label{fig:Figure6}
\end{figure}

\begin{figure}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/Figure7.pdf}
\end{center}
\caption{{\bf Histograms of the parameter estimates over 150 realisations}. The solid lines show the actual values of the parameters and the dotted lines show the means of the estimated parameters. (\textbf{A}) The histogram of the central excitatory connectivity kernel basis function parameter estimates, $\theta_0$. (\textbf{B}) The histogram of the surround inhibition connectivity kernel basis function parameter estimates, $\theta_1$. (\textbf{C}) The histogram of the longer range excitatory connectivity kernel basis function parameter estimates, $\theta_2$. (\textbf{D}) The histogram of the $\xi$ estimates where $\xi=1-\zeta T_s $ and $\zeta$ is the reciprocal of the synaptic time constant.} 
\label{fig:Figure7}
\end{figure}

\begin{figure}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/Figure9.pdf} 
\end{center}
\caption{{\bf Error in the field reconstruction}. The mean of average RMSE of the estimated field over 150 realisations (solid line) and the 95~\% confidence interval (grey region).} 
\label{fig:Figure9}
\end{figure}

\begin{figure*}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/Figure10.pdf}
\end{center}
\caption{{\bf Examples of the actual and estimated neural fields and the error}. (\textbf{A}) The actual field at a single time instant from the full neural field model that was used to generate the data. (\textbf{B}) The reconstructed field of the reduced model, showing the effect of the basis function decomposition on the frequency content of the estimated field. (\textbf{C}) The error between the true and reconstructed fields. The error shows high frequency behaviour.}
\label{fig:Figure10}
\end{figure*}

\begin{figure}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/Figure8_1.pdf}
\end{center}
\caption{{\bf Cross-section of connectivity kernel estimate}. The actual kernel is shown by the solid line, the mean kernel estimate (over 50 realisations) is shown by the dotted line and the 95~\% confidence interval is shown by the shaded grey region.}
\label{fig:Figure8}
\end{figure}

% \begin{figure}[!ht]
% \begin{center}
% \includegraphics{./Graph/pdf/OperatingRegions.pdf}
% \end{center}
% \caption{{\bf Operating Regions along the sigmoid}. The top graph is from the full model and the lower is from the reduced model.}
% \label{fig:OperatingRegions}
% \end{figure}

% \begin{figure*}[!ht]
% \begin{center}
% \includegraphics{./Graph/pdf/Figure11.pdf}
% \end{center}
% \caption{{\bf Hinton diagram of the RMSE of the kernel estimates}. Log of the RMSE of the kernel for different values of $\nu_{cm}$ and $\Delta_y$. Mean kernel estimates are obtained over 50 realisations in each case.}
% \label{fig:Figure11}
% \end{figure*}

\begin{figure*}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/RMSESigmoidParameters.pdf}
\end{center}
\caption{{\bf RMSE of different kernel estimates}. Log of the RMSE of the kernel for different values of firing threshold, $v_0$, and the slope of the sigmoid, $\varsigma$. Mean kernel estimates are obtained over 50 realisations in each case. White spaces indicate cases where the kernel can not be estimated due to numerical problems.}
\label{fig:Figure12}
\end{figure*}

\begin{figure*}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/sigmoidfunctions.pdf}
\end{center}
\caption{{\bf Sigmoid functions}. From bottom left corner to the right of the second second row of the previous figure.}
\label{fig:sigmoidfunctions}
\end{figure*}

% \begin{figure*}[!ht]
% \begin{center}
% \includegraphics{./Graph/pdf/Figure13.pdf}
% \end{center}
% \caption{{\bf RMSE of the kernel estimates}. RMSE of the kernel for different values of sensor width, $\sigma^2_m$, distance between adjacent sensors is $\Delta_y=1.5$. Mean kernel estimates are obtained over 50 realisations in each case.}
% \label{fig:Figure13}
% \end{figure*}

\begin{figure*}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/Figure13.pdf}
\end{center}
\caption{{\bf RMSE of the kernel estimates}. RMSE of the kernel for different values of sensor width, $\sigma_m$, distance between adjacent sensors is $\Delta_y=1.5$. Mean kernel estimates are obtained over 50 realisations in each case.}
\label{fig:Figure13}
\end{figure*}
\section*{Tables}

>>>>

\journal{Neuroimage}

\begin{document}

	\begin{frontmatter}
% Title must be 150 characters or less
\title{A Data-Driven Framework for Neural Field Modelling}

\author[DEEE,BEI,UE]{D. R. Freestone\corref{cor1}}
\ead{dfreestone@bionicear.org}
\author[US]{P. Aram} 
\author[UE,CU]{M. Dewar}
\author[UM]{K. Scerri}
\author[DEEE,BEI]{D. B. Grayden}
\author[US]{V. Kadirkamanathan}

\address[DEEE]{Department of Electrical and Electronic Engineering, University of Melbourne, Melbourne, VIC, Australia}
\address[BEI]{The Bionic Ear Institute, East Melbourne, VIC, Australia}
\address[UE]{Institute for Adaptive and Neural Computation, University of Edinburgh, Edinburgh, UK}
\address[US]{Department of Automatic Control and Systems Engineering, University of Sheffield, Sheffield, UK}
\address[CU]{Department of Applied Physics and Applied Mathematics, Columbia University, New York, NY, USA}
\address[UM]{Department of Systems and Control Engineering, University of Malta, Msida, MSD, Malta}

\begin{abstract}
This paper presents a framework for creating neural field models from electrophysiological data. The Wilson and Cowan or Amari style neural field equations are used to form a parametric model, where the parameters are estimated from data. To illustrate the estimation framework, data is generated using the neural field equations incorporating modelled sensors so a comparison can be made between estimated and true parameters. To facilitate state and parameter estimation, we introduce a method to reduce the continuum neural field model using a basis function decomposition to form a finite-dimensional state-space model. Spatial frequency analysis methods are introduced that systematically specify the basis function configuration required to capture the dominant characteristics of the neural field. The estimation procedure consists of a two-stage iterative algorithm incorporating the unscented Rauch-Tung-Striebel smoother for state estimation and a least squares algorithm for parameter estimation. The results show that it is theoretically possible to reconstruct the neural field and estimate intracortical connectivity structure and synaptic dynamics with the proposed framework. %The results also illustrate a systematic loss of high spatial frequency information as the cost incurred by the model reduction procedure. 
\end{abstract}

\begin{keyword}
	keyword1 \sep keyword2
%% keywords here, in the form: keyword \sep keyword

%% MSC codes here, in the form: \MSC code \sep code
%% or \MSC[2008] code \sep code (2000 is the default)

\end{keyword}

\end{frontmatter}

@@

	 \int_\Omega \boldsymbol{\psi}^{\top} (\mathbf{r}-\mathbf{r}')f(\boldsymbol{\phi}^{\top}(\mathbf{r}')\mathbf{x}_t)d\mathbf{r}' d\mathbf{r} \boldsymbol{\theta}  \nonumber \\
	& + \xi\mathbf{x}_t + \boldsymbol{\Gamma}^{-1} \int_\Omega{\boldsymbol{\phi}(\mathbf{r}) e_t(\mathbf{r}) \, d\mathbf{r}}.
\end{align}

@@

Histograms of the parameter estimates obtained over the 150 realisations are shown in Figure~\ref{fig:Figure7}. In each of the subplots of the figure the true parameter value is shown by the solid line and the mean parameter estimate is shown by the dashed line. The actual parameter values, the mean parameter estimates, standard deviations, and biases for this parameter combination are given as the first entry in Table~\ref{tab:MexicanHatKernelResults}, where other entries in the table show results from Experiment II. The biases show the percentage differences between the actual values and the mean parameter estimates. For this parameter combination the mean estimates are in good agreement (within 1 standard deviation) with the actual values of the connectivity kernel parameters. However, the time constant parameter, $\xi$, is outside the estimate distribution (with bias of $2.56\%$), with a small variance in parameter estimates. We attribute this small bias to the loss of the high frequency components in the smoother estimated field with respect to the actual field. These results thus indicate that choosing to limit the frequency content of the reconstructed neural field has only minor effects on the accuracy of the estimated kernel while allowing for significant gains in the computational demand.

The accuracy of the state estimates was evaluated by comparing the field reconstruction to the true field using the mean (over iterations) of the root mean square error (MRMSE) over space. The MRMSE of the field estimates over time for 150 realisations is shown in Figure~\ref{fig:Figure9} with the $95~\%$ confidence interval. At each time point, the error is consistently around $0.5$~mV, which is approximately $10.73~\%$ of the field voltage range ([-2.20, 2.46]~mV). The error in the reconstructed field is again mostly due to the loss of high spatial frequency information from the basis function decomposition as illustrated in Figure~\ref{fig:Figure10}. The figure shows a snapshot at a single time instant of the true neural field, the estimated field, and the corresponding error, which is the difference between the true and estimated neural fields. It can be seen in Figure~\ref{fig:Figure10}(B) that the estimated field contains the basic structure of the true field, shown in Figure~\ref{fig:Figure10}(A), but is missing the higher-frequency detail, which can be see in the error shown in Figure~\ref{fig:Figure10}(C).

@@

% The size of the squares in the figure show the magnitude of the log of the error. The squares are black if the log of the error is negative (the RMSE is $<1$) and the squares are white if the log of the error is positive (the RMSE is $>1$). The largest black squares indicate the best performance and the largest white squares indicate the worst performance for the parameter combinations (\dean{need some thing more quantitative}). 
From Figure~\ref{fig:Figure12} it can be seen that the estimation accuracy depends on the parameters of the sigmoid function. This is expected since by altering the parameters we are changing the nonlinearity and the saturation region of the sigmoid.

\ken{what I would write}

For example, when the sigmoid slope is very gentle, most of the neural activity falls in the central, nearly linear region of the sigmoid, thus having minor effects on the estimation procedure. On the other hand, sharp slopes result in most of the activity falling in the saturation region of the sigmoid with adverse effects on the estimation. Similar arguments can be made for variations to the sigmoid threshold.

\ken{what do you think?}

\dean{PLEASE HELP TO FILL THIS SECTION IN. ANYTHING WILL DO.}
..\dean{This next bit is CRAP (please help):} When the threshold is negative or the slope is very gentle, all of the activity is being passed through the sigmoid function. However, as the slope is increased or the threshold is increased, more of the field is within the saturating region of the activation function. This is illustrated in Figure~\ref{fig:sigmoidfunctions}, where the sigmoids from the bottom left corner through to near top right corner of Figure~\ref{fig:sigmoidfunctions} are plotted. This appears to have a detrimental effect on the estimation accuracy.

\subsubsection{Experiment IV: Results for Varying the Sensor Kernel Width}
The accuracy of the estimation algorithm was tested for different observation kernels by varying the width parameter, $\sigma_m$. Figure~\ref{fig:Figure13} shows the RMSE of the connectivity kernel estimates for the various observation kernel widths. As the theory presented earlier in Section~\ref{SpectralAnalysisSection} suggests, the observation kernel acts as a spatial filter.

\ken{tried to make it a little more straight forward}

When the observation kernel is too narrow, high frequency components in the neural field are sampled. These high frequency components result in aliasing errors in the estimated field thus affecting the kernel estimation accuracy. On the other hand, for the very wide kernels considered, important neural dynamics are not pick up by the estimated field due to the filtering out of the high frequency components by the sensors, resulting in poor accuracy in the kernel estimate. Thus, as indicated in Figure~\ref{fig:Figure13}, the best results are obtained when the sensor width is chosen to minimize both sources of errors.

\ken{that ok?}

When the observation kernel is too narrow, a small amount of high spatial frequency information is still present but not adequately sampled by the observation function, since the field is not exactly band-limited. Therefore, a small amount of aliasing occurs which results in the estimation error. When the observation kernel width is increased, the high spatial frequency information that is not adequately sampled is filtered more strongly by the observation kernel and the error decreases. However, as expected when the observation kernel becomes too large, important spatial frequency information is lost and the error becomes larger.

@@

\section{Conclusion}
\dean{I MOVED THIS FROM THE START OF THE DISCUSSION. ANY OBJECTIONS?} \ken{None} This paper has presented a framework for creating data-driven neural field models by developing a method for estimating intracortical connectivity and synaptic dynamics from electrophysiological data. In doing so, we have highlighted important aspects of the experiment design process, demonstrating the relationship between sensor bandwidth and the complexity of the field that can be inferred. We have derived a finite-order state-space representation of the continuum neural field equations that allows inference to be performed efficiently. Importantly, the mathematical formulation that we have presented scales well with the number of sensors. The major contribution of this framework is to formally bridge a gap between experimental and theoretical neuroscience in order to gain a better understanding of the mesoscopic dynamics of the brain.

\section{Acknowledgments}
Dean Freestone would like to thank Dr. Mark van Rossum for hosting him while visiting the University of Edinburgh where this work was undertaken and Dr. Matthias Hennig for useful discussions and feedback. This research was funded in part by the Australian Research Council (Linkage Project LP0560684). The Bionic Ear Institute acknowledges the support it receives from the Victorian State Government through the Operational Infrastructure Support Program. Dean Freestone would like to thank the University of Melbourne's Scholarships Office for the Postgraduate Overseas Research Experience Scholarship and the Harold Mitchell Foundation for a traveling scholarship for supporting this research. Parham Aram would also like to thank Andrew Hills for useful discussions and assistance with code and the generation of figures. The authors would also like to thank Dr. John Terry for insightful discussions and feedback.

\bibliographystyle{unsrt} 
\bibliography{BrainIDE}

\section*{Tables}

\begin{table*}[!ht]
\caption{
\bf{Notation}}
\begin{tabular}{|l|l|l|}
	\hline
	\textbf{Symbol} & \textbf{Quantity} & \textbf{Units}\\
	\hline
	\multicolumn{3}{|c|}{\emph{Domain and indices}}\\
	\hline
	$\Omega$ & spatial domain & n.a.\\
	$\mathbf{r}$ & spatial location & [mm, mm]\\
	$t$ & time & s\\
	\hline
	\multicolumn{3}{|c|}{\emph{Model}}\\
	\hline
    $y_t$ & observation & mV\\
    $v(\mathbf{r},t)$ & mean membrane potential & mV \\
	$g(\mathbf{r},t)$ & weighted firing rate function & spikes.s$^{-1}$\\
	$f(\mathbf{r},t)$ & firing rate function & spikes.s$^{-1}$\\
	$e(\mathbf{r},t)$ & field disturbance, with covariance function $\gamma(\mathbf r-\mathbf r')$ & mV\\
	$\boldsymbol\varepsilon_t$ & observation noise, with covariance matrix $\Sigma_\varepsilon$ & mV\\
	$m(\mathbf{r}_n,\mathbf{r}')$ & sensor kernel where $n$ is sensor index $n=1,..,n_y$ (width $\sigma_m^2$) & n.a. \\
	$w(\mathbf{r},\mathbf{r}')$ & spatial connectivity kernel & n.a.\\
	$h(t)$ & post-synaptic response kernel & mV\\
	$\zeta$, $\xi$ & inverse synaptic time constant \& time constant parameter & s$^{-1}$, n.a.\\
	$\eta(t)$ & Heaviside step function & n.a.\\
	$\delta(t)$ & Dirac-delta function & n.a.\\
	$D$ & temporal differential operator & n.a.\\
	\hline
	\multicolumn{3}{|c|}{\emph{Reduced Model}} \\
	\hline
   	$\boldsymbol\phi(\mathbf{r})$ & vector of $n_x$ Gaussian basis functions & n.a.\\
   	$\mathbf{x}_t$ & state vector at time $t$ & mV\\
   	$\boldsymbol{\psi}(\mathbf{r}-\mathbf{r}')$ & vector of $n_{\theta}$ connectivity kernel basis functions & n.a.\\
   	$\boldsymbol{\Psi}(\mathbf{r}')$ & decomposition connectivity matrix & n.a.\\
   	$\boldsymbol{\Gamma}$ & inner product of field basis functions & n.a.\\
   	$Q()$ & state transition function & n.a.\\
   	$\mathbf{e}_t$ & state disturbance vector, with covariance $\Sigma_e$ & mV\\
   	$\mathbf{C}$ & observation matrix & n.a. \\
	\hline
	\multicolumn{3}{|c|}{\emph{Frequency Analysis}} \\
	\hline
    % $V(\nu)$ & Fourier transform of the neural field & Hz\\
	$\boldsymbol{\nu}$, $\boldsymbol{\nu}_c$ & spatial frequency and spatial cutoff frequency & Hz \\
	$\sigma_{\nu}^2$ & variance of Fourier transformed Gaussian basis function & Hz$^2$\\
	\hline
	\multicolumn{3}{|c|}{\emph{Estimation}} \\
	\hline
	$\mathcal{X}$ & matrix of sigma vectors & mV\\
	$W$ & matrix of sigma vector weights (scaling parameter $\lambda$) & n.a.\\
   	$\hat{\mathbf{x}}_t^{f-}$, $\hat{\mathbf{x}}_t^f$ & forward prior and posterior state estimates & mV\\
   	$\hat{\mathbf{x}}_t^{b-}$, $\hat{\mathbf{x}}_t^{b}$ & backward prior and posterior state estimates & mV\\
 	$P^{f-}_t$, $P^f_t$  & forward prior and posterior covariance matrices & mV$^2$\\
   	$P^{b-}_t$, $P^b_t$ & backward prior and posterior covariance matrices & mV$^2$\\
	$M_t$& cross-covariance matrix & mV$^2$\\
	$\mathcal K_{t} $ & filter gain & n.a.\\
	$S_t$ & smoother gain & n.a.\\
	$q()$ & nonlinear state function (used in LS estimator) & n.a.\\
   	$\mathcal{\hat{W}}$& vector of parameter estimates (used in LS estimator) & n.a.\\	
	\hline
\end{tabular}
\begin{flushleft}The first column of the table gives the symbols for all the notation used in the paper. The second column provides a brief description of the quantity, and the third column provides the SI units where relevant.
\end{flushleft}
\label{tab:Notation}
\end{table*}
\clearpage
\newpage
\renewcommand{\arraystretch}{1.7}
\begin{table*}[!ht]
\caption{
\bf{Algorithm for the Unscented RTS Smoother}}
\begin{tabular}{|c|}\hline
\multicolumn{1}{|p{16cm}|}{\textbf{1.} Forward initialisation} \\ 
$\hat{\mathbf x}_0, \mathbf P_0$ \\
\hline
\multicolumn{1}{|p{16cm}|}{\textbf{2.} Forward iteration: for $t \in \left\lbrace 0,\cdots, T\right\rbrace $, calculate the sigma points $\mathcal X_{i,t}^f$ using equations \ref{eq:sigmapoints1}-\ref{eq:sigmapoints3} and propagate through equation~\ref{eq:QmatrixForSigmapoints}} \\
$\mathcal X_{i,t+1}^{f-}=Q(\mathcal X_{i,t}^f) \quad i=0, \dots, 2n_x$\\
\multicolumn{1}{|p{16cm}|}{Calculate the predicted state and the predicted covariance matrix} \\
$\hat{\mathbf x}_{t+1}^{f-}=\sum_{i=0}^{2n_x} W_i^{(m)}\mathcal X_{i,t+1}^{f-} $ \\
$\mathbf P_{t +1}^{f-}=\sum_{i=0}^{2n_x} W_i^{(c)}(\mathcal X_{i,t+1}^{f-}-\hat{\mathbf x}_{t +1}^{f-})(\mathcal X_{i,t+1}^{f-}-\hat{\mathbf x}_{t +1}^{f-})^\top+\boldsymbol \Sigma_e$ \\ 
\multicolumn{1}{|p{16cm}|}{Compute the filter gain, the filtered state and the filtered covariance matrix using the standard Kalman filter update equations} \\
$\mathcal K_{t+1}=\mathbf P_{t +1}^{f-}\mathbf C ^\top(\mathbf C \mathbf P_{t +1}^{f-}\mathbf C ^\top+\boldsymbol \Sigma_{\varepsilon})^{-1}$ \\
$\hat{\mathbf x}_{t+1}^{f}=\hat{\mathbf x}_{t+1}^{f-}+\mathcal K_{t+1}(\mathbf y_{t+1}-\mathbf C\hat{\mathbf x}_{t +1}^{f-})$ \\
$\mathbf P_{t+1}^f=(\mathbf I - \mathcal K_{t+1}\mathbf C)\mathbf P_{t +1}^{f-}$\\ 
\hline
\multicolumn{1}{|p{16cm}|}{\textbf{3.} Backward initialisation}\\
$\mathbf P_T^b= \mathbf P_T^f, \quad \hat{\mathbf x}^b_T= \hat{\mathbf x}^f_T$ \\
\hline
\multicolumn{1}{|p{16cm}|}{\textbf{4.} Backward iteration: for $t \in \left\lbrace T-1, \cdots, 0 \right\rbrace $ calculate the sigma points $\mathcal X_{i,t}^b$ and propagate them through equation \ref{eq:QmatrixForSigmapoints}}\\
$\mathcal X_{i,t+1}^{b-}=Q(\mathcal X_{i,t}^b) \quad i=0, \dots, 2n_x$\\
\multicolumn{1}{|p{16cm}|}{Calculate the predicted state and the predicted covariance matrix}\\
$\hat{\mathbf x}_{t+1}^{b-}=\sum_{i=0}^{2n_x} W_i^{(m)}\mathcal X_{i,t+1}^{b-}$\\
$\mathbf P_{t +1}^{b-}=\sum_{i=0}^{2n_x} W_i^{(c)}(\mathcal X_{i,t+1}^{b-}-\hat{\mathbf x}_{t +1}^{b-})(\mathcal X_{i,t+1}^{b-}-\hat{\mathbf x}_{t +1}^{b-})^\top+\boldsymbol \Sigma_e $\\
$\mathbf M_{t +1}=\sum_{i=0}^{2n_x} W_i^{(c)}(\mathcal X_{i,t}^{b-}-\hat{\mathbf x}_{t}^{f})(\mathcal X_{i,t+1}^{b-}-\hat{\mathbf x}_{t+1}^{b-})^\top$\\
\multicolumn{1}{|p{16cm}|}{Compute the smoother gain, the smoothed state and the smoothed covariance matrix}\\
$\mathbf S_t=\mathbf M_{t +1}\left[ \mathbf P_{t +1}^{b-}\right] ^{-1} $\\
$\hat{\mathbf x}_t^b=\hat{\mathbf x}_t^f+\mathbf S_t\left[\hat{\mathbf x}_{t+1}^{b}-\hat{\mathbf x}_{t+1}^{b-}\right]$\\
$\mathbf P_{t}^{b}=\mathbf P_{t}^{f}+\mathbf S_t\left[\mathbf P_{t+1}^{b}-\mathbf P_{t+1}^{b-} \right]\mathbf S_t^\top $\\
\hline
\end{tabular}
\begin{flushleft}This table shows the steps in the unscented Rauch-Tung-Striebel smoother algorithm. The steps are iterated 11 times for our state estimation procedure. The least squares algorithm is run after each iteration to update the parameter estimates.
\end{flushleft}
\label{tab:UKFAlgorithm}
\end{table*}
\clearpage
\newpage
\renewcommand{\arraystretch}{1}
\begin{table*}[!ht]
\caption{
\bf{Parameters}}
\begin{tabular}{|c|l|c|l|}
	\hline
	\textbf{Symbol} & \textbf{Quantity} & \textbf{Value} & \textbf{Units}\\
	\hline
	\multicolumn{4}{|c|}{\emph{Domain and indices}}\\
	\hline
	$\Delta$ & spatial discretisation step & 0.5 & mm \\
	$T_s$ & time step & $0.001$ & s\\
	$T$ & number of time steps used in estimation & 400 & n.a.\\
	\hline 
\multicolumn{4}{|c|}{\emph{Model}}\\
	\hline
	$\tau$ & synaptic time constant & 0.01 & s$^{-1}$ \\
% 	$f_{max}$ & maximal firing rate & 10 & spikes.s$^{-1}$\\
	$\varsigma$ & slope of sigmoidal activation function & 0.56 \cite{Wendling2005} & mV$^{-1}$\\
	$v_0$ & firing threshold & 1.8 \cite{Marreiros2008}& mV  \\
	$\boldsymbol{\theta}$ & vector of connectivity kernel parameters & $\left[\begin{array}{ccc}
	100 &-80 &5
	\end{array}
	\right]^{\top}$ & mV.spike$^{-1}$\\
	$\sigma_{\psi_{1}}, \sigma_{\psi_{2}}, \sigma_{\psi_{3}}$ & connectivity kernel width parameters & 1.8, 2.4, 6 & mm\\
